Self-Supervised Learning of Generalized Policies in adversarial domains

### 2021-06-30 doc/paper/paper.tex
problem with feature_generator
doc updated

\subsection{Nim}
\subsubsection{Description}
This is the Nim version used in~\cite{silver2020few}.In this domain there are two stacks (columns) of tokens. The number of actions available in each state depends on the number of tokens, a valid action involves removing one of the tokens. When a token is removed, that token and all the tokens above it are removed as well. When a token is removed, the turn is changed and the other player must remove a token as an action if there is one. The objective of the game is to leave the opponent with no actions available on his turn, i.e. without any token. The agent and the adversary has on action per turn.
It is interesting to note that in this game not all states are solvable if the opponent plays optimally. Knowing that the expert policy is to leave the opponent with both stacks with the same number of tokens, this state with both stacks leveled will not be solvable. Therefore, in order to solve it, we have made use of the $Bad(state)$ feature, which allows us not to have to define a good transition for each state, and as an alternative to label it as a bad state. The advantage is that if the initial state is solvable, bad states can always be avoided. However the policy is not complete for all states, in particular there will be no good transition for states with the two token stacks leveled, as we will see below, which we can say is the optimal representation of the game strategy. The winning tactic in this game is to level the two stacks of tokens by removing the token next to an empty cell and diagonally above another token, como se puede ver en~\cite{silver2020few}. The intuition behind this tactic is that if we follow it we can always reach a final situation where the opponent is forced to remove all the tokens from one pile, then we can remove the remaining tokens from the other pile.

@@

\subsection{Checkmate whith Queen}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the grid. In order to win the white pieces has to do a checkmate in one movement. The white pieces have the white-king and the white-queen, and black pieces have the black-king.

@@

\subsubsection{Task instance distribution}
To generate all the possible states from which the queen can checkmate the black king in one, in a board of $NxM$, we follow the following procedure. We generate all the possible checkmates that can be given to the black king if it is in the first row, then we apply three rotations to each board, 90, 180 and 270 degrees. To generate the checkmate with the black king in the first row we must take into account that for each black king position in this row there are three possible positions of the white king that allow a checkmate in one. In all these possible positions of the two kings we place the king in all the squares where it can checkmate the black king, and eliminate those where it would be giving a check. For the training, one hundred random boards of five by fifty are selected. To check that the policy is complete, it is tested on all possible, for each board size $NxM$ where both $N$ and $M$ are numbers no greater than twenty randomly selected.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  K  .  .  #  .  .  .  .  .  .  .  K  .  .  #
# .  .  .  .  .  .  .  .  .  k  #  .  .  .  .  .  .  .  .  Q  k  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  Q  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
\end{Verbatim}

\subsection{Checkmate with Rook}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the board. In order to win the white pieces has to do a checkmate in one movement. In white pieces have the white-king and the white-rook, and black pieces have the black-king.

>>>>

\subsection{Nim}
\subsubsection{Description}
This is the Nim version used in~\cite{silver2020few}.In this domain there are two stacks (columns) of tokens. The number of actions available in each state depends on the number of tokens, a valid action involves removing one of the tokens. When a token is removed, that token and all the tokens above it are removed as well. When a token is removed, the turn is changed and the other player must remove a token as an action if there is one. The objective of the game is to leave the opponent with no actions available on his turn, i.e. without any token. The agent and the adversary has one action per turn.
It is interesting to note that in this game not all states are solvable if the opponent plays optimally. Knowing that the expert policy is to leave the opponent with both stacks with the same number of tokens, this state with both stacks leveled will not be solvable. Therefore, in order to solve it, we have made use of the $Bad(state)$ feature, which allows us not to have to define a good transition for each state, and as an alternative to label it as a bad state. The advantage is that if the initial state is solvable, bad states can always be avoided. However the policy is not complete for all states, in particular there will be no good transition for states with the two token stacks leveled, as we will see below, which we can say is the optimal representation of the game strategy. The winning tactic in this game is to level the two stacks of tokens by removing the token next to an empty cell and diagonally above another token, como se puede ver en~\cite{silver2020few}. The intuition behind this tactic is that if we follow it we can always reach a final situation where the opponent is forced to remove all the tokens from one pile, then we can remove the remaining tokens from the other pile.

@@

\subsection{Checkmate in n with Queen}
\subsubsection{Description}

\subsubsection{Training instances}
\begin{Verbatim}[fontsize=\footnotesize]
  # .  k  .  .  K #
  # .  .  .  .  . #
  # .  .  .  .  . #
  # Q  .  .  .  . #
  # .  .  .  .  . #
\end{Verbatim}

\subsubsection{Features}
\begin{Verbatim}[fontsize=\footnotesize]
  (CM) Atom[checkmate] [k=1]
  (SM) Atom[stalemate] [k=1]
  (QbK) Atom[white_queen_between_kings] [k=1]
  (KE) Atom[black_king_closed_in_edge_by_white_queen] [k=1]
  (KD) Dist[cell-has-white_king;adjacent_cell;cell-has-black_king] [k=4]
  (BKhA) Bool[And(cell-attaked-by-black_king,Not(cell-attaked-by-white))] [k=4]
  (A) Num[And(cell-attaked-by-black_king,Not(cell-attaked-by-white_queen))] [k=4]
  () Bool[And(row-has-white_queen,Exists(adjacent_row,row-has-black_king)))] [k=6]
  (WQadjBK_col) Bool[And(col-has-white_queen,Exists(adjacent_col,col-has-black_king)))] [k=6]
  (WQadjBK) Bool[And(cell-has-white_queen,Exists(adjacent_cell,cell-has-black_king)))] [k=6]
  (WQadjWK) Bool[And(cell-has-white_queen,Exists(adjacent_cell,cell-has-white_king)))] [k=6]
\end{Verbatim}

\subsubsection{All Transition Features}
This features reason about how the value of the features can change in all possible successors.
\begin{Verbatim}[fontsize=\footnotesize]
  (CA) Bool[can(A, DEC)] [k=2]
  (CCM) bool[can(CM, INC)] [k=2]
\end{Verbatim}

\subsubsection{Feature Description}
$CM$. Checkmate\\
$SM$. Stalemate\\
$KE$. Black king is closed in an edge of the table by the white queen\\
$QbK$. White queen is somewhere between kings, rows or columns\\
$A$. The number of cells attaked by black king and not attacked by white queen\\
$BKhA$. Exist a cell attaked by Black king and no attacked by white pieces\\
$Qaw$. White queen is attaked and without protection, i.e. $WQadjBK AND WQadjWK$\\
$WQadjBK_line$. White queen is in an adjacent line to the black king, i.e. $WQadjBK_row OR WQadjWK_col$\\
$CA$. It is possible to decrease the number of cells attaked by black king and not attacked by white queen\\
$CCM$. It is possible to do checkmate in the current state

\subsubsection{Policy}
  \begin{Verbatim}[fontsize=\footnotesize]
   1. KE=0, CA>0 -> {A ↓}
   2. KE=0, CA=0 -> {A ·, KD ·}
   3. BKhA=0, WQadjBK_line>0 -> {KD ·, WQadjBK_line ·}
   4. CCM=0, KE>0 -> {KD ↓}
   5. CCM=1 -> {CM ↑}
  \end{Verbatim}

\subsubsection{Policy Implicits}
The next features has to be in that certain value after the transition for classifing the transition as good, taking into account the policy rules.
  \begin{Verbatim}[fontsize=\footnotesize]
   Qaw=0, Qbk>0, SM=0
  \end{Verbatim}

\subsubsection{Policy description}
Rules 1 and 2 cover the cases in which the king is not enclosed in an edge of the file. The first rule states that it is good to take transitions that decrease the number of squares available to the black king by using the queen. The second rule says that if the first rule is not applicable it is good to take transitions that do not alter the number of squares available to the black king, without moving the white king. The third rule refers to those cases where the black king had moved to its only free square, and classifies as good those transitions in which the queen is moved so that it is not stalemate and remains on the line adjacent to the black king. Rule four covers cases in which the black king is locked on an edge by the queen, and it is good to move the white king closer. finally, rule 5 says that if it is possible to checkmate it is good to do so. It should be noted that thanks to the implicit parts of the policy the king will never be left tied up and unprotected, stalemate will never occur and the white king will always be placed and remain in a line that is between the lines of the kings, whether rows or columns.

\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
  K.... # K.... # K.... # ..... # ..... # ..... # ..... # ..... # ..... #
  ..Q.. # ..... # ..... # ..K.. # ..K.. # ..K.. # ..... # ..... # ..... #
  ..... # ..... # ..... # ..... # ..... # ..... # ..K.. # ..K.. # ..K.. #
  ..... # Q.... # Q.... # Q.... # Q.... # Q.... # Q.... # Q.... # ...Q. #
  ....k # ....k # ...k. # ...k. # ..k.. # ...k. # ...k. # ....k # ....k #
\end{Verbatim}

\subsection{Checkmate in one with Queen}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the grid. In order to win the white pieces has to do a checkmate in one movement. The white pieces have the white-king and the white-queen, and black pieces have the black-king.

@@

\subsubsection{Task instance distribution}
To generate all the possible states from which the queen can checkmate the black king in one, in a board of $NxM$, we follow the following procedure. We generate all the possible checkmates that can be given to the black king if it is in the first row, then we apply three rotations to each board, 90, 180 and 270 degrees. To generate the checkmate with the black king in the first row we must take into account that for each black king position in this row there are three possible positions of the white king that allow a checkmate in one. In all these possible positions of the two kings we place the king in all the squares where it can checkmate the black king, and eliminate those where it would be giving a check. For the training, one hundred random boards of five by fifty are selected. To check that the policy is complete, it is tested on all possible, for each board size $NxM$ where both $N$ and $M$ are numbers no greater than twenty randomly selected.
v
\subsubsection{Example of Game}
\begin{Verbatim}[fontsize=\footnotesize]
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  K  .  .  #  .  .  .  .  .  .  .  K  .  .  #
# .  .  .  .  .  .  .  .  .  k  #  .  .  .  .  .  .  .  .  Q  k  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  Q  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
# .  .  .  .  .  .  .  .  .  .  #  .  .  .  .  .  .  .  .  .  .  #
\end{Verbatim}

\subsection{Checkmate in one with Rook}
\subsubsection{Description}
In this game the action space is equal to the number of cells in the board. In order to win the white pieces has to do a checkmate in one movement. In white pieces have the white-king and the white-rook, and black pieces have the black-king.
